---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 Assignment 1"
output: pdf_document
---

# Problem 1

Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`. Print only the result `vec_c`. Do not use the print command.

```{r}
vec_a <- c(2,4,6)
vec_b <- c(9,10,12)
vec_c <- vec_a + vec_b
vec_c
```

# Problem 2

Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a` and show the result. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?
```{r}
vec_d <- c(14,20)
vec_a <- vec_a + vec_d
vec_a
```
It added vector d to the first two values of vector a and made the third value a twenty. 
R gives me the message "longer object length is not a multiple of the shorter object length" thus explaining that it cannot add vector d to vector a correctly.

# Problem 3

Next add 5 to the vector `vec_a` and show the result. What is the result and what did R do? Why doesn't R give you a warning message similar to what you saw in the previous problem?
```{r}
vec_a + 5

```
R added 5 to each of the values. In the last problem the reason that it gave me an error was becuase it was the second vector was not a multiple of the first vector. Here there is only one value in the vector being added to the vector a so there are no issues. 


# Problem 4

Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function
```{r}
vect_seq <- seq(1,5)
vect_seq
```

b. Using the `a:b` shortcut.
```{r}
vect_15 <- 1:5 
vect_15
```

# Problem 5

Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the `seq()` function 
```{r}
vec_even <- seq(2, 20 , 2)
vec_even
```


b. Using the a:b shortcut and some subsequent algebra. 
```{r}
vec_shrt <- 1:10 *2
vec_shrt
```

# Problem 6

Generate a vector of 21 elements that are evenly spaced between 0 and 1 using the `seq()` command and name this vector `x`. Print the vector `x`.

```{r}
x <- seq(0,1, length.out = 21)
x
```

# Problem 7

Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
vec_rpt <- rep(c(2,4,8), 3)
vec_rpt
```


# Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for `rep()` to see all of the options that `rep()` will accept. In particular, look at the optional argument `each=`.
  
  
```{r}
vect_rpt2 <- rep(c(2,4,8), each = 4)
vect_rpt2
```
  
# Problem 9 

In this problem, we will work with the matrix 
$$ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]$$
    
a.  Follow the prompts below to create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands. Print `M`.
```{r}
M <- matrix(seq(2,30, 2), nrow= 3 , ncol=5, byrow = TRUE)
M
```

ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command. Print `M`
```{r}
M <- rbind(seq(2,10, 2), seq(12,20, 2), seq(22,30, 2))
M
```

b. Extract the second row out of `M`.
```{r}
M[2,]
```


c. Extract the element in the third row and second column of `M`.

```{r}
M[3,2]
```

# Problem 10 

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```
The both data frames search for the grades that are not NA but in different ways. The first uses a negatice sign before which and this makes sure that it does the opossite of what is asked inside the function. The second uses an ! to say not is.na, therfore making sure that the function finds the grades that are not NA.

# Problem 11 

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
```{r}
my.test <- list( x = c(4,5,6,7,8,9,10),
      y = c(34,35,41,40,45,47,51),
      slope = 2.82,
      p.value = 0.000131)
my.test
```

To show you have created your list correctly, type use `str()` to show the structure of the list.
```{r}
str(my.test)
```

b.  Extract the second element in the list.
```{r}
my.test[2]
```


c.  Extract the element named `p.value` from the list.

```{r}
my.test["p.value"]
```


