---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 S26 Assignment 5"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(mosaicData)
```

## Problem 1
Convert the following to date or date/time objects.

a. September 13, 2010.
```{r}
make_date( 2010, 9, 13)
```


b. Sept 13, 2010.

c. Sep 13, 2010.

d. S 13, 2010. Comment on the month abbreviation needs.
```{r}
mdy("Sept 13, 2010", "Sep 13, 2010", "S 13, 2010")
```
Abbreviation is need as there are many different months that share the same first letter. 

e. 07-Dec-1941.

f. 1-5-1998. Comment on why you might be wrong.
 Here i might be wrong because in the rest of the world the first number is the day and then the second is the month. from this format i have no way of knowing which data is correct. 
 
g. 21-5-1998. Comment on why you know you are correct.
There is no 21 month so this clearly is in day, month , year format. 
```{r}
dmy("07-Dec-1941", "1-5-1998", "21-5-1998")
```

h. 2020-May-5 10:30 am
```{r}
ymd_hm("2020-May-5 10:30 am")
```


i. 2020-May-5 10:30 am PDT (ex Seattle)

```{r}
ymd_hm("2020-May-5 10:30 am",tz = "America/Los_Angeles")
```

j. 2020-May-5 10:30 am AST (ex Puerto Rico)

```{r}
ymd_hm("2020-May-5 10:30 am",tz = "America/Puerto_rico")
```

## Problem 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following:

a. Calculate the date of your 64th birthday.
```{r}
birthday <- mdy("April, 29, 2005")

birthday + dyears(64)
```

b. Calculate your current age (in years). 
```{r}
current.age = interval(birthday,  lubridate::today() )

as.period(current.age) %>%
  year()

```


c. Using your result in part (b), calculate the date of your next birthday.
```{r warning=FALSE}



time_till_nextBirthday<- interval(current.age, dyears(21)) %>%
  as.period()

time_till_nextBirthday
```

d. The number of _days_ until your next birthday.
```{r warning=FALSE}

days.till <- interval(current.age, dyears(21)) %>%
  as.period(unit = "days") %>%
  day()
  

days.till
```

e. The number of _months_ and _days_ until your next birthday.
```{r warning=FALSE}
MD.till <- interval(current.age, dyears(21)) %>%
  as.period() %>%
  month()
  

MD.till
```


## Problem 3

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 

```{r warning=FALSE}

AZ <- mdy_hm( "May 8, 2015, 15:00" , tz = "US/Arizona")
AK <- with_tz(AZ,"Pacific/Auckland")

AK
```

## Problem 4

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 
```{r}
data("Births78")
```

```{r}
my.78 <- Births78

my.78 <- my.78  %>%
  select("date", 'births')
```

b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?
```{r}
ggplot(my.78,
       aes(x = date,
           y = births))+
  geom_point()
```

c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.
```{r}

my.78 <- rowwise(my.78) %>%
  mutate( dow = wday(date,label=TRUE, abbr=FALSE))

my.78 
```

d. Plot the data with the point color being determined by the day of the week variable.

```{r}
ggplot(my.78,
       aes(x = date,
           y = births,
           color = dow))+
  geom_point()
```









